---
title: "Figure1"
author: "Hao He"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{Figure1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE)
```

# Figure 1. The number of recombinations
a) Why don’t we try a single Sinaplot that includes all the results (GBRS, Muga, Megamuga, and Gigamuga) across all existing generations. 
Please use genotype-call based recombinations for GBRS. 
For this manuscript, I like to pick sigma=0.130 which gets us a consistent slope value around 18-20. 
I believe though I should re-run GBRS for Striatum data since I used sigma=0.120.

b) Gary wants a single scatterplot of array vs GBRS w.r.t # of recombinations. 
Each array type should be given a different color.

```{r Figure 1, message=FALSE, eval=TRUE, fig.height=20, fig.width=20}
library(sinaplot)
library(ggplot2)
library(plyr)

#load recombination results for three arrays
#DO_liver
#muga
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/muga275.recomb.RData")
muga275.recomb$array <- "MUGA"
muga275.recomb$array.sub <- "MUGA"
#megamuga
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/megamuga184.recomb.RData")
megamuga184.recomb$array <- "MEGAMUGA"
megamuga184.recomb$array.sub <- "MEGAMUGA"
#DO liver GBRS
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/DO_liver_recomb.GBRS.genotypes.RData")
gbrs.genotype.recomb <- recomb.GBRS.genotypes$`0.130`[recomb.GBRS.genotypes$`0.130`$Sample %in% c(as.character(muga275.recomb$Sample),as.character(megamuga184.recomb$Sample)),]
gbrs.genotype.recomb$array <- "GBRS"
gbrs.genotype.recomb$array.sub <- NULL
#GBRS.MUGA
gbrs.genotype.recomb[gbrs.genotype.recomb$Sample %in% c(as.character(muga275.recomb$Sample)),"array.sub"] <- "GBRS.MUGA"
#GBRS.MEGAMUGA
gbrs.genotype.recomb[gbrs.genotype.recomb$Sample %in% c(as.character(megamuga184.recomb$Sample)),"array.sub"] <- "GBRS.MEGAMUGA"

#DO_liver recombinations
DO_liver_recomb <- rbind(rbind(muga275.recomb,megamuga184.recomb),gbrs.genotype.recomb)
#merge with sample sheet in DO_liver
sample.sheet <- read.table("/projects/heh/HH/GBRS_array/data/DO_Liver/sample_sheet.tsv", header = TRUE)
DO_liver_recomb <- merge(DO_liver_recomb,sample.sheet, by = "Sample", all.x = TRUE)
DO_liver_recomb <- DO_liver_recomb[,c(1:6)]

#DO_Striatum
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/gigamuga358.recomb.RData")
gigamuga358.recomb$array <- "GIGAMUGA"
gigamuga358.recomb$array.sub <- "GIGAMUGA"
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/DO_Striatum_recomb.GBRS.genotypes.RData")
# gbrs.genotype.recomb <- gbrs.genotype.recomb[gbrs.genotype.recomb$Sample %in% gigamuga358.recomb$Sample,]
gbrs.genotype.recomb <- recomb.GBRS.genotypes$`0.130`[recomb.GBRS.genotypes$`0.130`$Sample %in% gigamuga358.recomb$Sample,]
gbrs.genotype.recomb$array <- "GBRS"
gbrs.genotype.recomb$array.sub <- "GBRS.GIGAMUGA"
#DO_striatum recombinations
DO_striatum_recomb <- rbind(gigamuga358.recomb,gbrs.genotype.recomb)
#merge with sample sheet in DO_Striatum
sample.sheet <- read.table("/projects/heh/HH/GBRS_array/data/DO_Striatum/sample_sheet.tsv", header = TRUE)
DO_striatum_recomb <- merge(DO_striatum_recomb,sample.sheet, by = "Sample", all.x = TRUE)
DO_striatum_recomb <- DO_striatum_recomb[,c(1:6)]

#all recomb results
recomb <- rbind(DO_liver_recomb,DO_striatum_recomb)
recomb$Gen <- as.factor(recomb$Gen)
recomb$array <- as.factor(recomb$array)
recomb$array.sub <- factor(recomb$array.sub, 
                           levels = c("GBRS.MUGA", "MUGA", 
                                      "GBRS.MEGAMUGA","MEGAMUGA",
                                      "GBRS.GIGAMUGA","GIGAMUGA"))

#get summary table for each generation per array
summary.tab <- ddply(recomb,.(array,Gen),
                           summarise,mean = mean(recomb),var = var(recomb), n= length(recomb))
#display summary table
DT::datatable(summary.tab,options = list(pageLength = 100, scrollY = "800px", scrollX = "20px"))

# for every individual sample recombination
recomb_ind <- merge(recomb[recomb$array == "GBRS",], recomb[recomb$array != "GBRS",], by = "Sample")
recomb_ind$array.sub.y <- as.factor(as.character(recomb_ind$array.sub.y))
#scatter plot
plot_recomb_ind <- ggplot(recomb_ind, aes(x=recomb.x, y=recomb.y)) + 
  geom_point(aes(shape = array.sub.y, color = array.sub.y), size = 4) + 
  scale_shape_manual(values=c(18,17,16)) +
  scale_color_manual(values=c("green3","blue","cyan")) +
  scale_fill_manual(values=c("green3","blue","cyan")) +
  theme(aspect.ratio=1) +
  xlab("Recombination in GBRS") +
  ylab("Recombination in array") +
  ggtitle("Recombination for each sample in GBRS vs array") + 
  scale_y_continuous(limits=c(100, 600)) +
  scale_x_continuous(limits=c(100, 600)) +
  geom_abline(slope=1, intercept=0) +
  theme(legend.title=element_blank(),legend.text=element_text(size=22), 
        text = element_text(size=22), axis.title=element_text(size=22)) +
  theme(plot.title = element_text(hjust=0.5, size = 22)) +
  guides(colour = guide_legend(override.aes = list(size=5)))

plot_recomb_ind

#Figure1_sinaplot
#color
# GBRS <- "red",
# MUGA <- "cyan",
# MEGAMUGA <- "blue",
# GIGAMUGA <- "green3"
#pdf(file = "Figure1_sinaplot_array.sub.pdf", width = 12, height = 10)
sinaplot(formula = recomb ~ array.sub, data = recomb, col = c("red","cyan","red","blue","red","green3"), pch = c(15,18,15,17,15,16), bty = "n",
         xlab = "",
         ylab = "Number of Recombinations",
         ylim=c(50, 600), cex.axis=1.5, cex.lab=1.5, cex = 1.5)
  legend("top", inset=.02, title="Array type",
   c("GBRS.MUGA", "MUGA", "GBRS.MEGAMUGA","MEGAMUGA","GBRS.GIGAMUGA","GIGAMUGA"), col=c("red","cyan","red","blue","red","green3"), pch = c(15,18,15,17,15,16), horiz=TRUE, cex=1.7)
#dev.off()

#pdf(file = "Figure1_sinaplot_array.pdf", width = 12, height = 10)
sinaplot(formula = recomb ~ array, data = recomb, col = c("red","green3","blue","cyan"), pch = c(15:18), bty = "n",
         xlab = "",
         ylab = "Number of Recombinations",
         ylim=c(50, 600), cex.axis=1.5, cex.lab=1.5, cex = 1.5)
  legend("top", inset=.02, title="Array type",
   c("GBRS","GIGAMUGA", "MEGAMUGA", "MUGA"), col=c("red","green3","blue","cyan"), pch = c(15:18), horiz=TRUE, cex=1.7)
#dev.off()

## sina plot
#pdf(file = "Figure1_scatterplot.pdf", width = 12, height = 10)
lm_eqn <- function(df){
    df$x <- as.numeric(as.character(df$x))
    m <- lm(y ~ x, df);
    eq <- substitute(italic(y) == a + b %.% italic(x), 
         list(a = format(coef(m)[1], digits = 2), 
              b = format(coef(m)[2], digits = 2)))
    as.character(as.expression(eq));                 
  }
recomb$Gen <- as.numeric(as.character(recomb$Gen))
#without slope
p2 <- ggplot(recomb, aes(x=Gen, y=recomb)) + 
  geom_jitter(aes(color=array, shape = array), size = 3, position=position_jitter(0.35))+
  scale_shape_manual(values=c(15:18)) +
  scale_x_continuous(breaks = c(4,5,NA,7,rep(NA,3),11,rep(NA,9),21:23)) +
  scale_color_manual(values=c("red","green3", "blue", "cyan")) +
  scale_fill_manual(values=c("red","green3", "blue", "cyan")) +
  labs(x = "DO Outbreeding Generations") +
  labs(y = "Number of Recombinations") +
  theme(legend.title=element_blank(), text = element_text(size=22), 
        axis.title=element_text(size=22,face="bold")) +
  guides(shape = guide_legend(override.aes = list(size = 5)))
p2
#with slope
p3 <- ggplot(recomb, aes(x=Gen, y=recomb)) + 
  geom_jitter(aes(color=array, shape = array), size = 3, position=position_jitter(0.35))+
  scale_shape_manual(values=c(15:18)) +
  scale_x_continuous(breaks = c(4,5,NA,7,rep(NA,3),11,rep(NA,9),21:23)) +
  geom_smooth(method = "lm", se=FALSE, color="black", aes(group=1), linetype="dotted") +
  geom_text(x = 11, y = 500, size = 14, 
            label = lm_eqn(df = data.frame(y = recomb$recomb, x = recomb$Gen)), parse = TRUE) +
  #scale_x_discrete(labels=c(as.character(4:23))) +
  scale_color_manual(values=c("red","green3", "blue", "cyan")) +
  scale_fill_manual(values=c("red","green3", "blue", "cyan")) +
  labs(x = "DO Outbreeding Generations") +
  labs(y = "Number of Recombinations") +
  theme(legend.title=element_blank(), text = element_text(size=22), 
        axis.title=element_text(size=22,face="bold")) +
  guides(shape = guide_legend(override.aes = list(size = 5)))
p3
#dev.off()
```

