---
title: "Figure2"
author: "Hao He"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{Figure2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE)
```

# Figure 2. Correlations
a) Let’s try a single histogram (or maybe density plot?) that shows all three array types. But let’s not filter out the unmatching samples (bars in low correlation regions) to show there exist considerable sample mixups across datasets.
b) Gary wanted to count how many times genotype calls agree/disagree between GBRS and arrays along 69k grid points. I am thinking this should be a table.

```{r Figure 2, message=FALSE, eval=TRUE, fig.height=20, fig.width=20}
library(sinaplot)
library(ggplot2)
library(plyr)
library(ggridges)

#load mismatch results to get the correlation
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/mismatch_muga275GBRS.RData")
muga.cor <- mismatch_muga275GBRS$`0.130`
muga.cor$array <- "MUGA"
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/mismatch_megamuga184GBRS.RData")
megamuga.cor <- mismatch_megamuga184GBRS$`0.130`
megamuga.cor$array <- "MEGAMUGA"
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/mismatch_gigamuga358GBRS.RData")
gigamuga.cor <- mismatch_gigamuga358GBRS
gigamuga.cor$array <- "GIGAMUGA"
#correlation results
gbrs.array.cor <- rbind(rbind(muga.cor,megamuga.cor),gigamuga.cor)
#histgram
#pdf(file = "Figure2_correlation_histogram.pdf", width = 12, height = 10)
p <- ggplot(gbrs.array.cor, aes(x=self.cor, fill = array)) +
  geom_histogram(alpha=0.70, position="identity",binwidth=0.005) +
  scale_color_manual(values=c("green3", "blue", "cyan")) +
  scale_fill_manual(values=c("green3", "blue", "cyan")) +
  labs(x = "Correlation between GBRS and Array") +
  labs(y = "Count") +
  #scale_x_discrete(limits=c(-0.02,1)) +
  scale_x_continuous(breaks = seq(0, 1.0, by = 0.15)) +
  theme(legend.title=element_blank(), text = element_text(size=22), 
        axis.title=element_text(size=18,face="bold"))
p
#dev.off()

gbrs.array.cor$array <- as.factor(gbrs.array.cor$array)
p2 <- ggplot(gbrs.array.cor, aes(x=self.cor, y=array, group = array)) +
  geom_density_ridges(aes(x = self.cor, fill = array), stat = "binline", binwidth = 0.005, scale = 0.95,alpha = .8) +
  scale_color_manual(values=c("green3", "blue", "cyan")) +
  scale_fill_manual(values=c("green3", "blue", "cyan")) +
  labs(x = "Correlation between GBRS and Array") +
  labs(y = "") +
  #scale_x_discrete(limits=c(-0.02,1)) +
  scale_x_continuous(breaks = seq(0, 1.0, by = 0.15)) +
  scale_y_discrete(
    expand = c(0.01, 0), name = "",
    labels = c("GIGAMUGA", "MEGAMUGA", "MUGA")
  ) +
  theme(legend.title=element_blank(), text = element_text(size=25), 
        axis.title=element_text(size=22,face="bold"))
p2

p3 <- ggplot(gbrs.array.cor, aes(x=self.cor, y=array, group = array)) +
  geom_density_ridges(aes(x = self.cor, fill = array), scale = 0.95, bandwidth = 0.01) +
  scale_color_manual(values=c("green3", "blue", "cyan")) +
  scale_fill_manual(values=c("green3", "blue", "cyan")) +
  labs(x = "Correlation between GBRS and Array") +
  labs(y = "") +
  #scale_x_discrete(limits=c(-0.02,1)) +
  scale_x_continuous(breaks = seq(0, 1.0, by = 0.15)) +
  scale_y_discrete(
    expand = c(0.01, 0), name = "",
    labels = c("GIGAMUGA", "MEGAMUGA", "MUGA")
  ) +
  theme(legend.title=element_blank(), text = element_text(size=25), 
        axis.title=element_text(size=22,face="bold"))
 p3
```

```{r Figure 2 autosome, message=FALSE, eval=TRUE, fig.height=20, fig.width=20}
#load mismatch results to get the correlation
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/mismatch_autochr_muga275GBRS.RData")
muga.cor <- mismatch_autochr_muga275GBRS$`0.130`
muga.cor$array <- "MUGA"
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/mismatch_autochr_megamuga184GBRS.RData")
megamuga.cor <- mismatch_autochr_megamuga184GBRS$`0.130`
megamuga.cor$array <- "MEGAMUGA"
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/mismatch_autochr_gigamuga358GBRS.RData")
gigamuga.cor <- mismatch_autochr_gigamuga358GBRS
gigamuga.cor$array <- "GIGAMUGA"
#correlation results
gbrs.array.cor <- rbind(rbind(muga.cor,megamuga.cor),gigamuga.cor)
#histgram
#pdf(file = "Figure2_correlation_on_autosome_histogram.pdf", width = 12, height = 10)
p <- ggplot(gbrs.array.cor, aes(x=self.cor, fill = array)) +
  geom_histogram(alpha=0.70, position="identity",binwidth=0.005) +
  scale_color_manual(values=c("green3", "blue", "cyan")) +
  scale_fill_manual(values=c("green3", "blue", "cyan")) +
  labs(x = "Correlation between GBRS and Array") +
  labs(y = "Count") +
  #scale_x_discrete(limits=c(-0.02,1)) +
  scale_x_continuous(breaks = seq(0, 1.0, by = 0.15)) +
  theme(legend.title=element_blank(), text = element_text(size=22), 
        axis.title=element_text(size=18,face="bold"))
p
#dev.off()

gbrs.array.cor$array <- as.factor(gbrs.array.cor$array)
p2 <- ggplot(gbrs.array.cor, aes(x=self.cor, y=array, group = array)) +
  geom_density_ridges(aes(x = self.cor, fill = array), stat = "binline", binwidth = 0.005, scale = 0.95,alpha = .8) +
  scale_color_manual(values=c("green3", "blue", "cyan")) +
  scale_fill_manual(values=c("green3", "blue", "cyan")) +
  labs(x = "Correlation between GBRS and Array on autosome markers") +
  labs(y = "") +
  #scale_x_discrete(limits=c(-0.02,1)) +
  scale_x_continuous(breaks = seq(0, 1.0, by = 0.15)) +
  scale_y_discrete(
    expand = c(0.01, 0), name = "",
    labels = c("GIGAMUGA", "MEGAMUGA", "MUGA")
  ) +
  theme(legend.title=element_blank(), text = element_text(size=25), 
        axis.title=element_text(size=22,face="bold"))
p2

p3 <- ggplot(gbrs.array.cor, aes(x=self.cor, y=array, group = array)) +
  geom_density_ridges(aes(x = self.cor, fill = array), scale = 0.95, bandwidth = 0.01) +
  scale_color_manual(values=c("green3", "blue", "cyan")) +
  scale_fill_manual(values=c("green3", "blue", "cyan")) +
  labs(x = "Correlation between GBRS and Array on autosome markers") +
  labs(y = "") +
  #scale_x_discrete(limits=c(-0.02,1)) +
  scale_x_continuous(breaks = seq(0, 1.0, by = 0.15)) +
  scale_y_discrete(
    expand = c(0.01, 0), name = "",
    labels = c("GIGAMUGA", "MEGAMUGA", "MUGA")
  ) +
  theme(legend.title=element_blank(), text = element_text(size=25), 
        axis.title=element_text(size=22,face="bold"))
p3
```

# Get agree table
```{r Agree table, message=FALSE, eval=FALSE}
# Library -----------------------------------------------------------------
# Load packages
library(qtl2)
library(tidyr)
library(dplyr)
library(data.table)
library(abind)

#MUGA -----------------------------------------------------------------
#load muga275.probs.69kchr  
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/data/qtl2/GENO/MG/muga275.probs.69kchr.RData")
#For each individual at each position, find the genotype with the maximum marginal probability.
muga275.g <- maxmarg(muga275.probs.69kchr,cores = 20, minprob = 0.5)
muga275.g <- do.call("abind",list(muga275.g,along = 2))

#GBRS275.geno.probs
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/data/qtl2/GENO/GBRS/GBRS275.geno.probs.RData")
attr(GBRS275.geno.probs, "crosstype") <- "do"
attr(GBRS275.geno.probs, "is_x_chr") <- structure(c(rep(FALSE,19),TRUE), names=1:20)
attr(GBRS275.geno.probs, "alleles") <- LETTERS[1:8]
attr(GBRS275.geno.probs, "alleleprobs") <- FALSE
attr(GBRS275.geno.probs, "class") <- c("calc_genoprob", "list")
gbrs275.g <- maxmarg(GBRS275.geno.probs$`0.130`,cores = 20, minprob = 0.5)
gbrs275.g <- do.call("abind",list(gbrs275.g,along = 2))

#agree and disagree in muga275
muga275.z <- muga275.g == gbrs275.g
muga275.z1 <- apply(muga275.z,1, function(x){length(which(x == TRUE))})
muga275.z2 <- apply(muga275.z,1, function(x){length(which(x == FALSE))})

#MEGAMUGA -----------------------------------------------------------------
#load megamuga184.probs.69kchr  
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/data/qtl2/GENO/MM/megamuga184.probs.69kchr.RData")
#For each individual at each position, find the genotype with the maximum marginal probability.
megamuga184.g <- maxmarg(megamuga184.probs.69kchr,cores = 20)
megamuga184.g <- do.call("abind",list(megamuga184.g,along = 2))

#GBRS184.geno.probs
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/data/qtl2/GENO/GBRS/GBRS184.geno.probs.RData")
attr(GBRS184.geno.probs, "crosstype") <- "do"
attr(GBRS184.geno.probs, "is_x_chr") <- structure(c(rep(FALSE,19),TRUE), names=1:20)
attr(GBRS184.geno.probs, "alleles") <- LETTERS[1:8]
attr(GBRS184.geno.probs, "alleleprobs") <- FALSE
attr(GBRS184.geno.probs, "class") <- c("calc_genoprob", "list")
gbrs184.g <- maxmarg(GBRS184.geno.probs$`0.130`,cores = 20)
gbrs184.g <- do.call("abind",list(gbrs184.g,along = 2))

#agree and disagree in megamuga184
megamuga184.z <- megamuga184.g == gbrs184.g
megamuga184.z1 <- apply(megamuga184.z,1, function(x){length(which(x == TRUE))})
megamuga184.z2 <- apply(megamuga184.z,1, function(x){length(which(x == FALSE))})

#GIGAMUGA -----------------------------------------------------------------
#load gigamuga358.probs.69kchr  
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/data/qtl2/GENO/GM/gigamuga358.probs.69kchr.RData")
#For each individual at each position, find the genotype with the maximum marginal probability.
gigamuga358.g <- maxmarg(gigamuga358.probs.69kchr,cores = 20)
gigamuga358.g <- do.call("abind",list(gigamuga358.g,along = 2))

#GBRS358.geno.probs
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/data/qtl2/GENO/GBRS/GBRS358.geno.probs.RData")
attr(GBRS358.geno.probs, "crosstype") <- "do"
attr(GBRS358.geno.probs, "is_x_chr") <- structure(c(rep(FALSE,19),TRUE), names=1:20)
attr(GBRS358.geno.probs, "alleles") <- LETTERS[1:8]
attr(GBRS358.geno.probs, "alleleprobs") <- FALSE
attr(GBRS358.geno.probs, "class") <- c("calc_genoprob", "list")
gbrs358.g <- maxmarg(GBRS358.geno.probs,cores = 20)
gbrs358.g <- do.call("abind",list(gbrs358.g,along = 2))

#agree and disagree in gigamuga358
gigamuga358.z <- gigamuga358.g == gbrs358.g
gigamuga358.z1 <- apply(gigamuga358.z,1, function(x){length(which(x == TRUE))})
gigamuga358.z2 <- apply(gigamuga358.z,1, function(x){length(which(x == FALSE))})


# Summary table for agree and disagree ------------------------------------
tab <- data.frame(Array = c("MUGA", "MEGAMUGA", "GIGAMUGA"),
                  
                  Agree = c( paste0(round(mean(muga275.z1),2), " ± ", round(sd(muga275.z1), 2)),
                             paste0(round(mean(megamuga184.z1),2), " ± ", round(sd(megamuga184.z1), 2)),
                             paste0(round(mean(gigamuga358.z1),2), " ± ", round(sd(gigamuga358.z1), 2))),
                  
                  Disagree = c( paste0(round(mean(muga275.z2),2), " ± ", round(sd(muga275.z2), 2)),
                             paste0(round(mean(megamuga184.z2),2), " ± ", round(sd(megamuga184.z2), 2)),
                             paste0(round(mean(gigamuga358.z2),2), " ± ", round(sd(gigamuga358.z2), 2))),
                  
                  total.percentage.agree = c( round(100*sum(muga275.z1)/c(dim(muga275.z)[1]*dim(muga275.z)[2]),3),
                                              round(100*sum(megamuga184.z1)/c(dim(megamuga184.z)[1]*dim(megamuga184.z)[2]),3),
                                              round(100*sum(gigamuga358.z1)/c(dim(gigamuga358.z)[1]*dim(gigamuga358.z)[2]),3)),
                  
                  total.percentage.disagree = c( round(100*sum(muga275.z2)/c(dim(muga275.z)[1]*dim(muga275.z)[2]),3),
                                                 round(100*sum(megamuga184.z2)/c(dim(megamuga184.z)[1]*dim(megamuga184.z)[2]),3),
                                                 round(100*sum(gigamuga358.z2)/c(dim(gigamuga358.z)[1]*dim(gigamuga358.z)[2]),3)))
write.csv(tab, file = "/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/Agree_table_three_array.csv")  
save(tab, file = "/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/Agree_table_three_array.RData")  
```

# show agree table
```{r, message=FALSE, eval=TRUE}
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/Agree_table_three_array.RData")                       
#display summary table
DT::datatable(tab,options = list(pageLength = 40, scrollY = "300px", scrollX = "40px"))
```


