---
title: "Figure1"
author: "Hao He"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
  toc: true
vignette: >
  %\VignetteIndexEntry{Figure1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-opts, include=FALSE}
knitr::opts_chunk$set(comment = "#",collapse = TRUE)
```

# Figure 1. The number of recombinations
a) Why don’t we try a single Sinaplot that includes all the results (GBRS, Muga, Megamuga, and Gigamuga) across all existing generations. 
Please use genotype-call based recombinations for GBRS. 
For this manuscript, I like to pick sigma=0.130 which gets us a consistent slope value around 18-20. 
I believe though I should re-run GBRS for Striatum data since I used sigma=0.120.

b) Gary wants a single scatterplot of array vs GBRS w.r.t # of recombinations. 
Each array type should be given a different color.

```{r Figure 1, message=FALSE, eval=TRUE, fig.height=20, fig.width=20}
library(sinaplot)
library(ggplot2)
library(plyr)

#load recombination results for three arrays
#DO_liver
#muga
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/muga275.recomb.RData")
muga275.recomb$array <- "MUGA"
muga275.recomb$array.sub <- "MUGA"
#megamuga
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/megamuga184.recomb.RData")
megamuga184.recomb$array <- "MEGAMUGA"
megamuga184.recomb$array.sub <- "MEGAMUGA"
#DO liver GBRS
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/DO_liver_recomb.GBRS.genotypes.RData")
gbrs.genotype.recomb <- recomb.GBRS.genotypes$`0.130`
gbrs.genotype.recomb$array <- "GBRS"
gbrs.genotype.recomb$array.sub <- NULL
#GBRS.MUGA
gbrs.genotype.recomb[gbrs.genotype.recomb$Sample %in% c(as.character(muga275.recomb$Sample)),"array.sub"] <- "GBRS.MUGA"
#GBRS.MEGAMUGA
gbrs.genotype.recomb[gbrs.genotype.recomb$Sample %in% c(as.character(megamuga184.recomb$Sample)),"array.sub"] <- "GBRS.MEGAMUGA"

#DO_liver recombinations
DO_liver_recomb <- rbind(rbind(muga275.recomb,megamuga184.recomb),gbrs.genotype.recomb)
#merge with sample sheet in DO_liver
sample.sheet <- read.table("/projects/heh/HH/GBRS_array/data/DO_Liver/sample_sheet.tsv", header = TRUE)
DO_liver_recomb <- merge(DO_liver_recomb,sample.sheet, by = "Sample", all.x = TRUE)
DO_liver_recomb <- DO_liver_recomb[,c(1:6)]

#DO_Striatum
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/gigamuga358.recomb.RData")
gigamuga358.recomb$array <- "GIGAMUGA"
gigamuga358.recomb$array.sub <- "GIGAMUGA"
load("/projects/heh/HH/GBRS_array/Workflowr_Array_GBRS/output/DO_Striatum_recomb.GBRS.genotypes.RData")
gbrs.genotype.recomb <- gbrs.genotype.recomb[gbrs.genotype.recomb$Sample %in% gigamuga358.recomb$Sample,]
gbrs.genotype.recomb <- recomb.GBRS.genotypes$`0.130`
gbrs.genotype.recomb$array <- "GBRS"
gbrs.genotype.recomb$array.sub <- "GBRS.GIGAMUGA"
#DO_striatum recombinations
DO_striatum_recomb <- rbind(gigamuga358.recomb,gbrs.genotype.recomb)
#merge with sample sheet in DO_Striatum
sample.sheet <- read.table("/projects/heh/HH/GBRS_array/data/DO_Striatum/sample_sheet.tsv", header = TRUE)
DO_striatum_recomb <- merge(DO_striatum_recomb,sample.sheet, by = "Sample", all.x = TRUE)
DO_striatum_recomb <- DO_striatum_recomb[,c(1:6)]

#all recomb results
recomb <- rbind(DO_liver_recomb,DO_striatum_recomb)
recomb$Gen <- as.factor(recomb$Gen)
recomb$array <- as.factor(recomb$array)
recomb$array.sub <- factor(recomb$array.sub, 
                           levels = c("GBRS.MUGA", "MUGA", 
                                      "GBRS.MEGAMUGA","MEGAMUGA",
                                      "GBRS.GIGAMUGA","GIGAMUGA"))

#get summary table for each generation per array
summary.tab <- ddply(recomb,.(array,Gen),
                           summarise,mean = mean(recomb),var = var(recomb), n= length(recomb))
#display summary table
DT::datatable(summary.tab,options = list(pageLength = 100, scrollY = "800px", scrollX = "20px"))

#Figure1_sinaplot
#color
# GBRS <- "red",
# MUGA <- "cyan",
# MEGAMUGA <- "blue",
# GIGAMUGA <- "green3"
#pdf(file = "Figure1_sinaplot_array.sub.pdf", width = 12, height = 10)
sinaplot(formula = recomb ~ array.sub, data = recomb, col = c("red","cyan","red","blue","red","green3"), pch = c(15,18,15,17,15,16), bty = "n",
         xlab = "",
         ylab = "Number of Recombinations",
         ylim=c(50, 600), cex.axis=1.5, cex.lab=1.5, cex = 1.2)
#dev.off()

#pdf(file = "Figure1_sinaplot_array.pdf", width = 12, height = 10)
sinaplot(formula = recomb ~ array, data = recomb, col = c("red","green3","blue","cyan"), pch = c(15:18), bty = "n",
         xlab = "DO Outbreeding Generations",
         ylab = "Number of Recombinations",
         ylim=c(50, 600), cex.axis=1.5, cex.lab=1.5, cex = 1.2)
#dev.off()

## sina plot
#pdf(file = "Figure1_scatterplot.pdf", width = 12, height = 10)
p2 <- ggplot(recomb, aes(x=Gen, y=recomb)) + 
  geom_jitter(aes(color=array, shape = array), size = 3, position=position_jitter(0.25))+
  scale_shape_manual(values=c(15:18))+
  scale_color_manual(values=c("red","green3", "blue", "cyan")) +
  scale_fill_manual(values=c("red","green3", "blue", "cyan")) +
  labs(x = "DO Outbreeding Generations") +
  labs(y = "Number of Recombinations") +
  theme(legend.title=element_blank(), text = element_text(size=25), 
        axis.title=element_text(size=20,face="bold"))
p2
#dev.off()
```

